\LARGE \bf Informed sampling by Markov Chain Monte Carlo methods

### 2017-02-10 paper.tex
reordering

Define the dynamics of a robot as 
\begin{equation}
\label{eq:kinematics}
\dot{x}(t) = f( x(t) , u(t) ), 
\end{equation}
in which $ x \in \calX \subseteq \mathbb{R}^n $ is the state of the robot, and $ u \in \calU \subseteq \mathbb{R}^m $ is the control input.
The dynamics determines how a state trajectory $ \gamma $ is like given duration $ T $, state $ x(t) $ and input $ u(t) $.
There exists a free trajectory space $ \Pi_{free} $ that contains all feasible trajectories $ \gamma \in \Gamma_{\rm free} $.
It is determined by the system dynamics,$ \Cfree $ and $ \Ufree $, in which $ \Cfree \subseteq \calX $ is a free state space considering the obstacles in a workspace, and $ \Ufree \subseteq \calU $ is a free control space considering the bounds of control input. 
The cost of a trajectory $ \gamma $ is defined as the accumulated cost $ c( x(t), u(t) ) $ along the path, which is 
\begin{equation}
\label{eq:path_cost}
c(\gamma) = \int_0^{T} c( x(t), u(t) ) dt.
\end{equation}
An optimal kinodynamic path-planning problem that finds a kinodynamic path minimizing the cost $ \pi^* = \min_{ \pi \in \Gamma_{free} } c( \gamma ) $.

Informed sampling-based planning algorithms (e.g. Informed RRT*~\cite{GSB14}, BIT*~\cite{GSB15}) samples new states only from an \emph{informed set}.
An informed set is a subspace $ \Cinf \subseteq \calX $   that is possible to provide better solution than the current best $  \Cinf = \{ x \in \calX \mid c ( \gamma^*(x) ) < \cbest \} $~\cite{GSB14}, in which $ \gamma^*(x) $ means the optimal path from $ x_s $ to $ x_g $ constrained to pass through $ x $ without considering collision checking, and $ \cbest $ is the cost of the current best found.
The current best $ \cbest $ constrains eligible $ x $ indirectly by an implicit $ \gamma^*(x) $.
In an optimal kinodynamic path-planning problem, more constraints are considered in defining an informed set, which include the system dynamic constraint and the boundary of control input.
If we consider a manifold that is defined by an informed set $ \Cinf $, the informed sampling becomes how to sample on this manifold in a high-dimensional space.

@@

See Fig. ~\ref{fig:sample_dimt} for an example motion of one joint through the $\left(q,\dot{q}\right)$ space.

@@

From these equations it shows a fundamental difference of a kinodynamic state space compared to a geometric state space. Because of this the minimum time path can vary in shape depending on the start and goal state. The consequence of this is that the shape of the informed subset can change dramatically depending on the start and goal state of the problem definition. Because of this, finding a closed form solution to sampling in the informed space is not feasible.

@@

\begin{figure}[tb]
  \centering
  	\includegraphics[width=0.5\textwidth]{fig/kino_space.png}
  \caption{
    \captionstyle
  	Side view of the cost surface for the case of minimal-time double integrator. The side view showcases the discontinuous nature of the cost surface.
  	}
   	\label{fig:cost_surface_1d}
	\vspace{-5.5mm}
\end{figure}

See Fig. ~\ref{fig:cost_surface_1d} for a side view of the cost surface for one degree of freedom ($\left(x, \dot{x}\right)$). One can see a key property of the cost surface that cause serious concerns is the discontinuous nature of the surface. As mentioned before, finding the minimum time problem amounts to solving a quadratic equation. Because the nature of this equation changes with the start and end goal, there is a curve in the input space that amounts to a switching surface that represents the boundary for the decision of the sign of $a_1$. This surface causes a discontinuous cost surface as shown in Fig. ~\ref{fig:cost_surface_1d}.

@@

%\dy{I am not sure what we want to put into here, but I am going to continue on the discussion of the algorithm and talk about some of the probability stuff.}

@@

In order to take advantage of these powerful methods, we propose to reformulate the cost surface as a probability density. Specifically, we consider a Boltzmann distribution in order to formulate the cost function using flexible energy based methods. Because Monte Carlo methods only consider the ratio of the probabilities, we do not need to consider the problem of estimating the intractable partition function.

\os{
	Roughly speaking, HMC views the system as an energy function for which Hamiltonian dynamics are simulated in fixed time steps.
	Thus, to sample from a desired distribution~$P$ using HMC, we need to convert the distribution to an energy function and vice-versa.
	This is done using the concept of a canonical distribution from statistical mechanics~\cite{N11}. 
	Given some energy function,~$E$, the canonical distribution over states has the Boltzman probability distribution $P = \frac{1}{Z}\exp (-E/T)$, where $Z$ is a normalizing constant and $T$ is the temperature of the system.
	%Viewing this the opposite way, if we are interested in some distribution with density function $P$, we can obtain it as a canonical distribution with $T = 1$ by setting $E = -\log P - \log Z$, where $Z$ is any convenient positive constant	
}

\os{
The energy function's definition provides the flexibility to specify desired or more probable states (by presenting lower energies) without the need to use a formal and restrictive model. 
Specifically, in our setting we want to approximate 
a uniform distribution over points inside the informed set~\Cinf.
Note that simulating Hamiltonian dynamics requires that the energy function defined is differentiable.
}

\os{
Our energy for any given time $t$, 
denoted by~$E\left(t,\Cbest \right)$, 
is a function of $t$ and the current best cost~$\Cbest$.
}

In order to ensure that the span the set of all possible states

@@

\section{Results}

@@

>>>>

Define the dynamics of a robot as 
\begin{equation}
\label{eq:kinematics}
\dot{\bm{x}}(t) = f( \bm{x}(t) , \bm{u}(t) ), 
\end{equation}
in which $ \bm{x} \in \mathcal{X} \subseteq \mathbb{R}^n $ is the state of the robot, and $ \bm{u} \in \mathcal{U} \subseteq \mathbb{R}^m $ is the control input.
The dynamics determines a state trajectory $ \pi $, given a duration $ T $, an initial state $ \bm{x}(0) $ and an initial input $ \bm{u}(0) $.
The cost of a trajectory $ \pi $ is defined as the accumulated cost along the path, which is 
\begin{equation}
\label{eq:path_cost}
c(\pi) = \int_0^{T} c( x(t), u(t) ) dt.
\end{equation}
Consider the obstacles in a workspace, we have $ \mathcal{X}_{free} \subseteq \mathcal{X} $.
Consider the bounds of the control input, we have $ \mathcal{U}_{free} \subseteq \mathcal{U} $.
The dynamics of the robot in Equation~\eqref{eq:kinematics}, free state space $  \mathcal{X}_{free} $ and free control input space $ \mathcal{U}_{free} $ determines a free trajectory space $ \Pi_{free} $, which contains all feasible trajectories $ \pi \in \Pi_{free} $.
An optimal kinodynamic path-planning problem that finds a kinodynamic path minimizing the cost $ \pi^* = \max_{ \pi \in \Pi_{free} } c( \pi ) $ subject to constraints $ \bm{x} (t) \in \mathcal{X}_{free}  $ and $ \bm{u} (t) \in \mathcal{U}_{free} $.

Informed sampling-based planning algorithms (e.g. Informed RRT*~\cite{GSB14}, BIT*~\cite{GSB15}) samples new states only from an \emph{informed set}.
An informed set is a subspace $ \Cinf \subseteq \calX $   that is possible to provide better solution than the current best $  \Cinf = \{ x \in \calX \mid c ( \pi^*(x) ) < c_{best} \} $~\cite{GSB14}, in which $ \pi^*(x) $ means the optimal path from $ x_{init} $ to $ x_{goal} $ constrained to pass through $ x $ and $ c_{best} $ is the cost of the current best found.
The current best $ c_{best} $ constrains eligible $ x $ indirectly by an implicit $ \pi^*(x) $.
In an optimal kinodynamic path-planning problem, more constraints are considered in defining an informed set, which include the system dynamic constraint and the boundary of control input.
If we consider a manifold that is defined by an informed set $ X_f $, the informed sampling becomes how to sample on this manifold in a high-dimensional space.

@@

\ignore{
See Fig. ~\ref{fig:sample_dimt} for an example motion of one joint through the $\left(q,\dot{q}\right)$ space.

@@

From these equations it shows a fundamental difference of a kinodynamic state space compared to a geometric state space. Because of this the minimum time path can vary in shape depending on the start and goal state. The consequence of this is that the shape of the informed subset can change dramatically depending on the start and goal state of the problem definition. Because of this, finding a closed form solution to sampling in the informed space is not feasible.
}

@@

\begin{figure}[tb]
  \centering
  	\includegraphics[width=0.5\textwidth]{fig/kino_space.png}
  \caption{
    \captionstyle
  	Side view of the cost surface for the case of minimal-time double integrator. The side view showcases the discontinuous nature of the cost surface.
  	}
   	\label{fig:cost_surface_1d}
%	\vspace{-5.5mm}
\end{figure}

See Fig.~\ref{fig:cost_surface_1d} for a side view of the cost surface for one degree of freedom ($\left(x, \dot{x}\right)$). One can see a key property of the cost surface that cause serious concerns is the discontinuous nature of the surface. As mentioned before, finding the minimum time problem amounts to solving a quadratic equation. Because the nature of this equation changes with the start and end goal, there is a curve in the input space that amounts to a switching surface that represents the boundary for the decision of the sign of $a_1$. This surface causes a discontinuous cost surface as shown in Fig. ~\ref{fig:cost_surface_1d}.

\section{MCMC-based Informed Sampling}
	Roughly speaking, HMC views the system as an energy function for which Hamiltonian dynamics are simulated in fixed time steps.
	Thus, to sample from a desired distribution~$P$ using HMC, we need to convert the distribution to an energy function and vice-versa.
	This is done using the concept of a canonical distribution from statistical mechanics~\cite{N11}. 
	Given some energy function,~$E$, the canonical distribution over states has the Boltzman probability distribution $P = \frac{1}{Z}\exp (-E/T)$, where $Z$ is a normalizing constant and $T$ is the temperature of the system.
	%Viewing this the opposite way, if we are interested in some distribution with density function $P$, we can obtain it as a canonical distribution with $T = 1$ by setting $E = -\log P - \log Z$, where $Z$ is any convenient positive constant

The energy function's definition provides the flexibility to specify desired or more probable states (by presenting lower energies) without the need to use a formal and restrictive model. 
Specifically, in our setting we want to approximate 
a uniform distribution over points inside the informed set~\Cinf.

Our energy for any given time $t$, 
denoted by~$E\left(t,\Cbest \right)$, 
is a function of $t$ and the current best cost~$\Cbest$.
To consider only points in the informed set, all with equal energy, one may consider taking 
$$
E\left(t,\Cbest \right) = 
\begin{cases}
    1,	& \text{if } t \leq \Cbest\\        
    0,  	& \text{otherwise}.
\end{cases}
$$
However, simulating Hamiltonian dynamics requires that the energy function defined is differentiable.
This function has a discontinuity at $t = \Cbest$ 
and contains no gradient information for $t \leq \Cbest$.
Thus, we define our energy function\footnote{In practice we add another term to the energy function which we discuss in our experiments (see Sec.~\ref{sec:eval}). We omit this term here for clarity of exposition.} as  
\begin{equation}
\label{eq:energy}
E\left( t, \Cbest \right) 
= 
E_{\text{grad}} \left( t \right) + 
E_{\text{informed}} \left( t,\Cbest \right).
\end{equation}
Here, $E_{\text{grad}}$ is a user-defined function that
controls the ``flatness`` of the distribution.
Intuitively, this term biases the distribution towards states with smaller cost and balances between the uniformity of the distribution and the gradient information provided to the HMC algorithm.
%
%In our setting we experimented with using 
%$\log\left(1 + \log\left(t\right)\right)$ and $\tanh\left(t\right)$. 
%
The term 
$E_{\text{informed}} \left( t,\Cbest \right) 
= \beta_{\text{informed}} \cdot \text{sigm}\left(t_0 - t\right)
$
penalizes states outside the informed set~$\Cinf$ using a tight sigmoid function~$\text{sigm}$ scaled by some constant~$\beta_{\text{informed}}$.
The sigmoid can be seen as a differentiable function smoothing the boundary of~$\Cinf$.

\subsection{Implementation details}
In our implementation we add another term to our energy function (Eq.~\ref{eq:energy}) relating to system parameters that induce known bounds on states.
Specifically, our robot has predefined joint limits. 
The flexibility of the energy function allows us to incorporate these limits  to penalize states \emph{within} the informed set that are close to, or violate these limits.

Specifically, we add to Eq.~\ref{eq:energy} the term 
$
E_{\text{limits}}\left(\mathbf{x}\right) = \beta_{\text{limits}} \cdot ||\mathbf{x}_{\text{limits}} - \mathbf{x}||_2
$
where $||\mathbf{x}_{\text{limits}} - \mathbf{x}||_2$ represents the L2 distance between the current state $\mathbf{x}$ and the state limits~$\mathbf{x}_{\text{limits}}$ and~$\beta_{\text{limits}}$ is some constant controlling how much to penalize states close to the limits.

@@

@@

@@

\section{Evaluation}
\label{sec:eval}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Old text that may be re-used
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ignore{
In order to take advantage of these powerful methods, we propose to reformulate the cost surface as a probability density. Specifically, we consider a Boltzmann distribution in order to formulate the cost function using flexible energy based methods. Because Monte Carlo methods only consider the ratio of the probabilities, we do not need to consider the problem of estimating the intractable partition function.
